A thesis draft

### 2018-11-16 res/sections/08-Background.tex
Fix typos

\section{Software-defined networking}
The SDN is an emerging approach to cloud computing that allows to configure
network in a programmatically way, possibly enhancing monitoring and
performances. The main sake of SDN is to the packet forwarding process,
\emph{data plane}, from the routing one, the so called \emph{control plane}.
The latter component serves as the centralized network intelligence and could
consists of one or more controllers. It has some drawbacks in teorms of
security, scalability and elasticity, because of the centralized architecture.

\subsection{SDN stack}
\begin{figure}[t]
 \centering
 \includegraphics[scale=1.3]{sdn_architecture}
 \caption[SDN architecture schema]{SDN architecture schema. Is possible to 
denote the three layer logic: the Application plane, the Controller plane and 
the Data plane.}
 \label{chap:background:img:sdn_architecture}
\end{figure}
As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
architecture can be splitted in three different layers, called planes, based on 
their functionalities.

@@

The \emph{Application plane} role is to set the behaviour of the SDN controllers 
and manage its hardware resources and data through APIs, often called 
\emph{northbound interfaces} (NBI).
These interfaces provide an additional abstraction layer that high-level 
applications can consume, enabling the possibility to use a variety of 
different verdors' controllers in the same way.
Analysing the SDN applications is possible to note two defined layers, one that 
is responsible to manage the whole logic, and another one dedicated to the 
communication part, that in particular uses the previsously defined NBIs.
These applications can also build an abstracted view of the network, using
controllers information, in order to make the traffic analysis easier and to 
simplify the decision making process.

@@

\emph{SDN controllers}, are logically centralized components that
receive requests from the mentioned SDN applications and either translate them 
in a fashion that can be understood from the lower layer\todo{who?}, or collect 
data needed by the application itself. 
Due to the variety of tasks a controller has to accomplish, its logical 
suddivision is achieved with a three-layer architecture, in which every layer 
exposes it's own interface, that can be consequently used by the upper 
layers, up to the \emph{NBI agents}.
NBI agent manages the communications with SDN applications, exposing interfaces
that usually provide an abstract view of the network and a set of
the controller functionalities.
In particular, the middle layer represents the controller inner-logic, which 
process the requests coming from the upper layer (aka the third layer), that 
has the specific task to communicate with the remaining network compoents, in 
other words the Control to Data-Plane Interface and/or the already 
mentioned southbound interface.

@@

In Figure~\ref{chap:background:img:openflow_protocol} it is possible to see the
how switches works in the OpenFlow protocol. Basically, OpenFlow swiches own a
set of tables to organized packet control and forwarding. These tables can be
updated by higher-level application, usually an SDN application or controller,
taking advatage of OpenFlow channels (called OF channels).

@@

\subsection{Network Virtualization}
%https://s3.amazonaws.com/academia.edu.documents/29016225/cs-2008-25.pdf?AWSAcce
%ssKeyId=AKIAIWOWYYGZ2Y53UL3A&Expires=1541352965&Signature=On6XJ5daSAmnfJMnNf41f
%BNHFgg%3D&response-content-disposition=inline%3B%20filename%3DA_survey_of_netwo
%r k _virtualization.pdf
Virtualization, in the networking context, refers to the act of combining
hardware, software and functionalities into a virtual network. This abstraction
runs on the top of a physical network in a hypervisor, decupling the logical
network from the harwdare one. Network virtualization allow multiple,
heterogeneous networks to cohabit, increasing the flexibility and manageability.
%http://www.csc.villanova.edu/~nadi/csc8580/S11/DeeptiNune.pdf
%https://www.sciencedirect.com/science/article/abs/pii/S1389128609003387
The idea of network virtualization came from the decoupling of the role of the
traditional ISP into two different entities: who provides the infrastucture
(InPs) and who provides the services(SPs). In that context it is possible to
aggregate resources, even from different InPs, creating a virtual network,
offering an end-to-end service. Most known instances of NV are \emph{VPNs},
dedicated networks that allows the communication among multiple sites, using
private and secure tunnels, relying on the public network such as Internet.
Other examples of netvork virtualization are \emph{overlay networks}. They are
logical networks, build on the top of an existing physical network, tipically
implemented at the application layer.

\subsection{Network Function Virtualization}
The \emph{Network Function Virtualization} (NFV) concept refers to the use of 
virtualization tecnologies to virtualize network nodes functions. VNFs may run 
different processes and softwares, on the top of servers, switches, storage 
devices or on cloud computing infrastuctures, that substitute custom hardware. 
That approach worked well in the past, because of the need to follow rigorous 
standards, providing at the same time high quality, but it led to long product 
cycles and high specialized proprietary hardware. New Internet senarios, that 
include IoT and the increasing number of services to spread multimedia data 
(Skype, Netflix, Spotify, ...), requires more flexible functions and faster 
network deployments. The NFV technology is exploited in a framework, organized
at least by three components:
\begin{enumerate}
 \item a software that performs the network function that must be deployed,
 that is the VNF;
 \item NFV Infrastructure (NFVI) that is composed by all hardware and software 
components that builds the environment where the VNFs are deployed;
 \item NFV Management And Orchestration (NFV-MANO) architectural framework that
 is all the functional blocks, data repositories used by these blocks,
 reference points and interfaces through which these functional blocks exchange
 information for the purpose of managing and orchestrating NFVI and VNFs.
\end{enumerate}

>>>>

\section{Software-defined networking}
The SDN is an emerging approach to cloud computing that allows to configure
network in a programmatically way, possibly enhancing monitoring and
performances. The main sake of SDN is to the packet forwarding process,
\emph{data plane}, from the routing one, the so called \emph{control plane}.
The latter component serves as the centralized network intelligence and could
consists of one or more controllers. It has some drawbacks in terms of
security, scalability and elasticity, because of the centralized architecture.

\subsection{SDN stack}
\begin{figure}[t]
 \centering
 \includegraphics[scale=1.3]{sdn_architecture}
 \caption[SDN architecture schema]{SDN architecture schema. Is possible to 
denote the three layer logic: the Application plane, the Controller plane and 
the Data plane.}
 \label{chap:background:img:sdn_architecture}
\end{figure}
As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
architecture can be split in three different layers, called planes, based on 
their functionalities.

@@

The \emph{Application plane} role is to set the behaviour of the SDN controllers 
and manage its hardware resources and data through APIs, often called 
\emph{northbound interfaces} (NBI).
These interfaces provide an additional abstraction layer that high-level 
applications can consume, enabling the possibility to use a variety of 
different vendors' controllers in the same way.
Analyzing the SDN applications is possible to note two defined layers, one that 
is responsible to manage the whole logic, and another one dedicated to the 
communication part, that in particular uses the previously defined NBIs.
These applications can also build an abstracted view of the network, using
controllers information, in order to make the traffic analysis easier and to 
simplify the decision making process.

@@

\emph{SDN controllers}, are logically centralized components that
receive requests from the mentioned SDN applications and either translate them 
in a fashion that can be understood from the lower layer\todo{who?}, or collect 
data needed by the application itself. 
Due to the variety of tasks a controller has to accomplish, its logical 
subdivision is achieved with a three-layer architecture, in which every layer 
exposes it's own interface, that can be consequently used by the upper 
layers, up to the \emph{NBI agents}.
NBI agent manages the communications with SDN applications, exposing interfaces
that usually provide an abstract view of the network and a set of
the controller functionalities.
In particular, the middle layer represents the controller inner-logic, which 
process the requests coming from the upper layer (aka the third layer), that 
has the specific task to communicate with the remaining network components, in 
other words the Control to Data-Plane Interface and/or the already 
mentioned southbound interface.

@@

In Figure~\ref{chap:background:img:openflow_protocol} it is possible to see the
how switches works in the OpenFlow protocol. Basically, OpenFlow switches own a
set of tables to organized packet control and forwarding. These tables can be
updated by higher-level application, usually an SDN application or controller,
taking advantage of OpenFlow channels (called OF channels).

@@

\subsection{Network Virtualization}
%https://s3.amazonaws.com/academia.edu.documents/29016225/cs-2008-25.pdf?AWSAcce
%ssKeyId=AKIAIWOWYYGZ2Y53UL3A&Expires=1541352965&Signature=On6XJ5daSAmnfJMnNf41f
%BNHFgg%3D&response-content-disposition=inline%3B%20filename%3DA_survey_of_netwo
%r k _virtualization.pdf
Virtualization, in the networking context, refers to the act of combining
hardware, software and functionalities into a virtual network. This abstraction
runs on the top of a physical network in a hyper-visor, decoupling the logical
network from the hardware one. Network virtualization allow multiple,
heterogeneous networks to cohabit, increasing the flexibility and manageability.
%http://www.csc.villanova.edu/~nadi/csc8580/S11/DeeptiNune.pdf
%https://www.sciencedirect.com/science/article/abs/pii/S1389128609003387
The idea of network virtualization came from the decoupling of the role of the
traditional ISP into two different entities: who provides the infrastructure
(InPs) and who provides the services(SPs). In that context it is possible to
aggregate resources, even from different InPs, creating a virtual network,
offering an end-to-end service. Most known instances of NV are \emph{VPNs},
dedicated networks that allows the communication among multiple sites, using
private and secure tunnels, relying on the public network such as Internet.
Other examples of network virtualization are \emph{overlay networks}. They are
logical networks, build on the top of an existing physical network, typically
implemented at the application layer.

\subsection{Network Function Virtualization}
The \emph{Network Function Virtualization} (NFV) concept refers to the use of 
virtualization technologies to virtualize network nodes functions. VNFs may run 
different processes and software, on the top of servers, switches, storage 
devices or on cloud computing infrastructures, that substitute custom hardware. 
That approach worked well in the past, because of the need to follow rigorous 
standards, providing at the same time high quality, but it led to long product 
cycles and high specialized proprietary hardware. New Internet scenarios, that 
include IoT and the increasing number of services to spread multimedia data 
(Skype, Netflix, Spotify, \dots), requires more flexible functions and faster 
network deployments. The NFV technology is exploited in a framework, organized
at least by three components:
\begin{enumerate}
 \item a software that performs the network function that must be deployed,
 that is the VNF;
 \item NFV Infrastructure (NFVI) that is composed by all hardware and software 
components that builds the environment where the VNFs are deployed;
 \item NFV Management And Orchestration (NFV-MANO) architectural framework that
 is all the functional blocks, data repositories used by these blocks,
 reference points and interfaces through which these functional blocks exchange
 information for the purpose of managing and orchestrating NFVI and VNFs.
\end{enumerate}
